Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-12 WSSSPE2_report/WSSSPE2_review.tex
light editing in 7.1.2-3 and comments

This group started by discussing who should be incentivized, thinking of two categories of people:  
those in science (who could be incentivized to do better, more shareable, more sustainable work), and
those in industry but interested in science (who could be incentivized to contribute to science.)  John Towns made the point that we are not yet clear enough on exactly what we want to incentivize, suggesting that we need to have a clearer picture of ``good computational work" and what sort of contributions are truly generative for science.

The discussion of incentivizing those in science acknowledged that the publications system was far from perfect for incentivizing good software work. Nonetheless discussion focused on bringing software people into publications. Two main suggestions: the first is to focus on end-users of software and encourage them to cite software the ``right" way. James Howison suggested that his research showed that few projects were making a formal request for citation (but that authors weren't necessarily following those suggestions anyway); he suggested making access to the software conditional on a license that requires citation.  Others found this ``too confrontational" and preferred to concentrate on making it easier to do the right thing.  The group was more enthusiastic about ``forcing" PIs to include their ``software people" on publications, although there were few ideas on how exactly to do this.  Another technique mentioned was that when scientific software projects are hosted in organizations like Apache, the scientific contributors can benefit from building their reputations, perhaps yielding job offers that they can use to negotiate better job and career packages.

The discussion on incentivizing those outside science focused on accessing the well of affection that those working in software have for scientific research. How can the interest and skills of this group be marshaled towards sustainable contributions?  The is evidence that the migration of scientific software projects to the Apache Foundation has created opportunities for those not employed in the scientific center to contribute to projects initiated for scientists (especially where there is cross-over with industry needs, such as provenance and workflow).

The group centered on a discussion of developing ``software prizes" arguing that while it is hard to ``mint" other new sources of reputation, prizes are possible without getting too many others on board. The prize criteria can form a template for describing what we mean by scientific contributions made through software, particularly focusing on building active communities, not only writing great code. Prizes were identified as something worthy of more discussion, prompting a later break-out group, discussed in Section \ref{sec:prizes}.

@@

In the remaining discussion sessions, the group chose to split into three subgroups to discuss
a version of these topics: citation ecosystems,
taxonomy of contributorship/guidelines for software for tenure review, and prizes.  The subgroups were asked to clearly identify
\begin{itemize}
\item the problem to be solved
\item steps towards a solution
\end{itemize}

>>>>

This group started by discussing who should be incentivized, thinking of two categories of people:  
those in science (who could be incentivized to do better, more shareable, more sustainable work), and
those in industry but interested in science (who could be incentivized to contribute to science.)  It was pointed out that we are not yet clear enough on exactly what we want to incentivize, suggesting that we need to have a clearer picture of ``good computational work" and what sort of contributions are truly generative for science.

The discussion of incentivizing those in science acknowledged that the publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications. There were two main suggestions.
The first is to focus on end users of software and encourage them to cite software the ``right" way. \katznote{what is the second?}
James Howison suggested that his research showed that few projects were making a formal request for citation (but that authors weren't necessarily following those suggestions anyway) \katznote{add cite?}; he suggested making access to the software conditional on a license that requires citation.
Others found this ``too confrontational" and preferred to concentrate on making it easier to do the right thing.
The group was more enthusiastic about ``forcing" PIs to include their ``software people" on publications, although there were few ideas on how exactly to do this.
Another technique mentioned was that when scientific software projects are hosted in organizations like Apache, the scientific contributors can benefit from building their reputations, perhaps yielding job offers that they can use to negotiate better job and career packages.

The discussion on incentivizing those outside science focused on accessing the well of affection that those working in software have for scientific research.
How can the interest and skills of this group be marshaled towards sustainable contributions?
The is evidence that the migration of scientific software projects to the Apache Foundation has created opportunities for those not employed in the scientific center to contribute to projects initiated by scientists (especially where there is cross-over with industry needs, such as provenance and workflow).

The group also discussed developing ``software prizes" arguing that while it is hard to ``mint" other new sources of reputation, prizes are possible without getting too many others on board.
The prize criteria can form a template for describing what we mean by scientific contributions made through software, particularly focusing on building active communities, not only writing great code.

@@

In the remaining discussion sessions, the group chose to split into three subgroups to discuss
a version of these topics: {\em citation ecosystems},
{\em taxonomy of contributorship/guidelines for software for tenure review}, and {\em prizes}.  The subgroups were asked to clearly identify
\begin{itemize}
\item the problem to be solved
\item steps towards a solution
\end{itemize}
